Evaluation of classical correlation functions from 2/3D images on CPU and GPU architectures: introducing CorrelationFunctions.jl

### 2022-10-07 paper-correlations.tex
Scan approach (1/N)

The aim of this paper is to establish a general open-source solution allowing to
compute all classical correlation functions. This solution should be well
documented with all computational algorithms explained, it should also leverage
on recent advances in programming high-intensity computations and utilize both
CPU and GPU power. Our answer in the form of \code{CorrelationFunctions.jl}
package allows to compute numerous CFs easily on any operating system. The
manuscript describing the package is organized as follows: in \cref{sec:math} we
introduce a reader to the most common and useful correlation functions,
\cref{sec:map} presents all algorithmic details for computing full CFs maps (map
method, full 2-point statistics for a given image). In \cref{sec:scan} we
provide all algorithmic details for CFs computation in directions (scanning
method). In \cref{sec:verification} we verify the all computations based on
known analytical solutions. \cref{sec:efficiency} deals with computational
efficiency of our code, we compare computational times for 2D and 3D images of
different sizes for scanning and map methods as evaluated using either CPU or
GPU. In \cref{sec:examples} some application examples are provided, including
CFs evaluation for multi-phase materials. Finally, \cref{sec:summary} summarizes
all results and outlines possible future improvements.

@@

One of basic correlation functions is a lineal path function which is defined as
follows for homogenous media:
\begin{definition}
  Lineal path function $L_2^{(i)}(\mathbf{r})$ equals to probability that a
  vector $\mathbf{r}$ lies wholly in phase $i$ when randomly thrown into the
  sample.
\end{definition}
If a sample is isotropic a vector $\mathbf{r}$ can be replaced with its length
$r$ and the definition becomes:
\begin{definition}
  Lineal path function (for isotropic media) $L_2^{(i)}(r)$ equals to
  probability that a line segment of length $r$ lies wholly in phase $i$ when
  randomly thrown into the sample.
\end{definition}

Another function of big importance is a two-point correlation function $S_2$. A
definition for homogenous media is as follows:
\begin{definition}
  Two point function $S_2^{(i)}(\mathbf{r})$ equals to probability that both ends
  of a vector $\mathbf{r}$ lie in phase $i$ when the vector is randomly thrown
  into the sample.
\end{definition}
For isotopic media a vector $\mathbf{r}$ can be again reduced to a scalar value
$r$ (i.e. a correlation function depends only on length of a vector and not on
its direction.) Mathematically this definition can be expressed with the
following equation:
\begin{equation}
  S_2^{(i)}(\mathbf{r}) = \langle I^{(i)}(\mathbf{x}) I^{(i)}(\mathbf{x} +
  \mathbf{r}) \rangle
  \label{eq:s2-def}
\end{equation}
where $I^{(i)}$ is an indicator function for a set of all points belonging to
the phase $i$ and $\langle \cdots \rangle$ is an average over all xs.

A function closely related to $S_2$ is phase cross-correlation function
$\rho_{ij}$ which is defined with the following equation:
\begin{equation}
  \rho_{ij}(\mathbf{r}) = \langle I^{(i)}(\mathbf{x}) I^{(j)}(\mathbf{x} +
  \mathbf{r}) \rangle
  \label{eq:cross-def}
\end{equation}

The next function is a cluster function $C_2$.
\begin{definition}
  Cluster function $C_2^{(i)}(\mathbf{r})$ equals to probability that both
  ends of a vector $\mathbf{r}$ lie in the same cluster when the vector is
  randomly thrown into the sample. A cluster is a set of points of the same
  phase $i$ where any to points can be connected by a path lying entirely in
  that phase.
\end{definition}

The next two functions describe surface of a sample. These functions are
surface-surface and surface-void functions. For homogenous media their
definitions are:
\begin{definition}
  Surface-surface function $F_{ss}^{(i)}(\mathbf{r})$ equals to probability that
  both ends of a vector $\mathbf{r}$ touch the interface of phase $i$ when the
  vector is randomly thrown into the sample.
\end{definition}
and
\begin{definition}
  Surface-void function $F_{sv}^{(i)}(\mathbf{r})$ equals to probability that
  one end of a vector $\mathbf{r}$ touches the interface of phase $i$ and the
  other end touches the void phase when the vector is randomly thrown into the
  sample.
\end{definition}
If we introduce an interface indicator function
\begin{equation*}
  M^{(i)}(\mathbf{r}) = | \nabla I^{(i)}(\mathbf{r}) |
\end{equation*}
we can define these functions mathematically as
\begin{align}
  F_{ss}^{(i)}(\mathbf{r}) &= \langle M^{(i)}(\mathbf{x}) M^{(i)}(\mathbf{x} +
  \mathbf{r}) \rangle \label{eq:fss-def} \\ 
  F_{sv}^{(i)}(\mathbf{r}) &= \langle M^{(i)}(\mathbf{x}) I^{(void)}(\mathbf{x}
  + \mathbf{r}) \rangle \label{eq:fsv-def}
\end{align}
As always, there are versions of these functions applicable to scalar argument
for isotropic media.

@@

Cluster function can be calculated in the similar manner. Firstly, we apply an
indicator function to an input $A$: $A^{(i)} \xleftarrow{I^{(i)}} A$. When we
label connected components in $A^{(i)}$ using one of many popular algorithms for
this purpose \textcolor{red}{cite something}. This algorithm must take into
account if the function is calculated in periodic mode or not. Then $C_2$
function is a sum of $S_2$ functions calculated for all clusters
separately. Algorithmic complexity of this implementation is $O(MC \log M)$
where $M$ is the number of elements in the input and $C$ is the number of
clusters. The algorithm is as follows:
\begin{algorithmic}[1]
  \Procedure{$C_2$}{$A, phase, periodic$}
  \State $A^{(phase)} \gets I^{(phase)} (A)$
  \State $CC \gets label\_components(A^{(phase)}, periodic)$
  \State $N \gets maximum(CC)$ \Comment Number of connected components
  \State \textbf{return} $\sum_{n=1}^N S_2(CC, n, periodic)$
  \EndProcedure
\end{algorithmic}

@@

\section{The scan approach.}
\label{sec:scan}
In this approach, the input array is cut in parallel slices (``scanned'') along
each requested direction. Each slice is then analyzed independently as
one-dimensional system and the result is accumulated into \code{CorrelationData}
structure. In the following sections we provide the algorithms for computation
of the functions given in \cref{sec:math}.

Generally a correlation function $F(\bm{r})$ is defined as a probability
$p(\bm{r})$ of event $A$ which can occur when a line segment of length
$|\bm{r}|$ and parallel to $\bm{r}$ is randomly dropped into the input
array. Examples of $A$ include an event that the whole segment lies in some
phase $i$ or that it touches an interface between phases with both of its ends
and so on.

The probability $p(\bm{r})$ can be expressed as
\begin{equation}\label{eq:probability}
  p(\bm{r}) = \frac{\text{Number of successful trials $S(\bm{r})$}}
  {\text{Total number of trials $N(\bm{r})$}}
\end{equation}

In the scan approach we fix one direction $\bm{e}$ and accumulate the values of
$S(\bm{r})$ and $N(\bm{r})$ so that $\bm{r} \parallel \bm{e}$ into two arrays
\code{success} and \code{total} respectively. We then calculate \code{success}
and \code{total} for all directions and construct a \code{CorrelationData}
structure which holds two dictionaries with key-value pairs
\code{(direction, success)} and \code{(direction, total)}. With this information
it's easy to calculate a correlation function for each combination of
correlation length and direction.

\subsection{Lineal-path function}
Lineal-path function $L_2^{(i)}(\bm{r})$ is defined as the probability that a
line segment with the length $|\bm{r}|$ and paralel to $\bm{r}$ lies wholly in
the phase $i$ when randomly thrown into the input array. This function for a
one-dimensional array is computed using the algorithm described in
\apref{linpathalg}. A multidimensional arrays is firstly cut in slices and
\code{l2} function is calculated for each slice returning \code{success} and
\code{total} arrays. These arrays are then accumulated as described in
\cref{sec:scan} for final calculation of $L_2^{(i)}(\bm{r})$ function.

\subsection{Two-point function}
\label{sec:s2}
Two-point correlation function $S_2^{(i)}(\bm{r})$ is defined as the probability
that both ends of a line segment with the length $|\bm{r}|$ and paralel to
$\bm{r}$ lie in the phase $i$ when the segment is randomly thrown into the input
array. This can be expressed using the following formula:
\begin{equation*}
  S_2^{(i)}(\bm{r}) = \langle I^{(i)}(\bm{x}) I^{(i)}(\bm{x} + \bm{r}) \rangle
\end{equation*}
where $I^{(i)}(\bm{x})$ is an indicator function for the phase $i$:
\begin{equation*}
I^{(i)}(\bm{x}) = \left\{
\begin{array}{ll}
  1 & \quad \bm{x} \in \text{phase $i$} \\
  0 & \quad \text{otherwise}
\end{array}
\right.
\end{equation*}

Two-point function is so heaviliy used internally in the package so we propose a
bit more general definition for it:
\begin{equation}
  S_2^{\chi}(\bm{r}) = \langle \chi(\bm{x}, \bm{x} + \bm{r}) \rangle
  \label{eq:s2gen}
\end{equation}
where $\chi(\bm{x_1}, \bm{x_2})$ can be understood as some arbitrary indicator
(or membership) function on $\mathbb{R}^{2n}$ where $n$ is dimensionality of the
input array. Vectors $\bm{x_1}$ and $\bm{x_2}$ are the coordinates of the both
ends of a line segment thrown into the input array.

If $\chi(\bm{x_1}, \bm{x_2})$ can be expressed as
$\chi(\bm{x_1}, \bm{x_2}) = \chi_1(\bm{x_1})\chi_2(\bm{x_2})$ (what we call
\code{SeparableIndicator} in our package) we use the following algorithm to
calculate the array \code{success} for one-dimensional slice $a$: \footnote{From
here on $\odot$ denotes concatenation and $:$ denotes consing (joining an
element in front of a list).}
\begin{algorithmic}[1]
  \Procedure{s2sep}{$a, \chi_1, \chi_2, periodic$}
  \If{$!periodic$}
    \State $a \gets a \odot [0 \dots 0]$ \Comment{Pad with zeros to twice the
      length of $a$.}
  \EndIf
  \State $a_1 \gets \chi_1(a)$ \Comment{apply $\chi_1$ element-wise to $a$}
  \State $a_2 \gets \chi_2(a)$ \Comment{apply $\chi_2$ element-wise to $a$}
  \State $f_1 \gets rfft(a_1)$ \Comment{$rfft$ is fast Fourier transform for
    real-valued data}
  \State $f_2 \gets rfft(a_2)$
  \State $f_{S_2} \gets f_1 \cdot \overline{f_2}$ \Comment{multiply element-wise}
  \State $success \gets irfft(f_{S_2})$ \Comment{$irfft$ is inverse of $rfft$}
  \State \textbf{return} $success$
  \EndProcedure
\end{algorithmic}

If $\chi(\bm{x})$ cannot be expressed in this way (what we call
\code{InseparableIndicator}) the algorithm is slightly different:
\begin{algorithmic}[1]
  \Procedure{s2insep}{$a, \chi, periodic$}
    \State $len \gets length(a)$
    \If{$periodic$}
      \State $a \gets a \odot a$
    \EndIf
    \State $success \gets zeros(len)$
    \For{$shift \gets 0,len-1$}
      \State $a_2 \gets a[shift \dots]$
      \Comment{Take elements with index bigger than $shift$.}
      \State $success[shift] \gets sum(\chi(a, a_2))$
      \Comment{Apply $\chi$ element-wise and take a sum.}
    \EndFor
    \State \textbf{return} $success$
  \EndProcedure
\end{algorithmic}

The array \code{total} is obtained in the same way as in $L_2$ function because
it depends only on correlation lengths and the length of the input array.

The efficiency of \code{s2sep} is $O(n\log n)$ and efficiency of
\code{s2insep} is $O(n^2)$ where $n$ is the length of the array.

\subsection{Cluster function}
The cluster correlation function $C_2^{(i)}(\bm{r})$ is defined as the
probability that both ends of a line segment with the length $|\bm{r}|$ and
paralel to $\bm{r}$ lie in the same cluster of the phase $i$ when the segment is
randomly thrown into the input array.

For computation of cluster function we use \code{label\_components} function
from \code{Image.jl} package which extracts connected comonents (clusters) from the
binary array. Each cluster gets its own unique integer identificator. The
identificator $0$ is reserved for the void phase.

The algorithm for calculation of the array \code{success} for
$C_2^{(i)}(\bm{r})$ for one-dimensional slice $a$ is straightforward:
\begin{algorithmic}[1]
  \Procedure{c2}{$a, phase, periodic$}
  \State $a_1 \gets x = phase$ \Comment{$a$ is compared with $phase$
    element-wise}
  \State $c \gets label\_components(a_1)$
  \State $\chi(x,y) \gets (x = y \ne 0)$
  \State \textbf{return} $s2insep(c, \chi, periodic)$
  \EndProcedure
\end{algorithmic}

\subsection{Surface-surface and surface-void functions}
The surface-surface correlation function $F_{ss}^{(i)}(\bm{r})$ is defined as
the probability that both ends of a line segment with the length $|\bm{r}|$ and
paralel to $\bm{r}$ touch the interface of the phase $i$ when the segment is
randomly thrown into the input array. This can be expressed using the following
formula:
\begin{equation*}
  F_{ss}^{(i)}(\bm{r}) = \langle |\nabla I^{(i)}(\bm{x})| |\nabla I^{(i)}(\bm{x}
  + \bm{r})| \rangle
\end{equation*}
where $I^{(i)}(\bm{x})$ is an indicator function for the phase $i$.

The algorithm for calculation of the array \code{success} for
$F_{ss}^{(i)}(\bm{r})$ for one-dimensional slice $a$is also straightforward:
\begin{algorithmic}[1]
  \Procedure{surfsurf}{$a, phase, periodic$}
    \State $a_1 \gets a = phase$
    \State $edge \gets \text{Detect edges in $a_1$}$
    \State \textbf{return} $s2sep(edge, E, E, periodic)$
    \Comment{$E(x) = x$}
  \EndProcedure
\end{algorithmic}

The surface-void correlation function $F_{sv}^{(i)}(\bm{r})$is defined as
follows:
\begin{equation*}
  F_{sv}^{(i)}(\bm{r}) = \langle |\nabla I^{(i)}(\bm{x})| I^{(void)}(\bm{x}
  + \bm{r}) \rangle
\end{equation*}

The algorithm for calculation of the array \code{success} for
$F_{sv}^{(i)}(\bm{r})$ for one-dimensional slice $a$ is very similar to the
previous one:
\begin{algorithmic}[1]
  \Procedure{surfvoid}{$a, phase, periodic$}
    \State $a_1 \gets a = phase$
    \State $edge \gets \text{Detect edges in $a_1$}$
    \State $indices \gets \text{Array of all indices in $a$}$
    \State $\chi_1(x) \gets edge[x]$
    \State $\chi_2(x) \gets a[x] = 0$ \Comment{$0$ is a value reserved for the
      void phase}
    \State \textbf{return} $s2sep(indices, \chi_1, \chi_2, periodic)$
  \EndProcedure
\end{algorithmic}

The edges are extracted by calculating $|S(a)|$ where $S$ is the Sobel
operator. \textbf{TODO: Cite our first paper}. Algorithmic efficiency of these
algorithms are $O(n\log n)$ because it's implemented using \code{s2sep} and
filtering to extract edges is $O(n)$.

\subsection{Pore size function}
The pore size correlation function $P(r)dr$ is defined as the probability that a
randomly chosen point in the input array lies at a distance between
$r$ and $r + dr$ from the nearest point on the pore-solid interface.

As you can see this function cannot be obtained from one-dimensional slices. We
calculate it using distance transform. For each point $\bm{x}$ distance
transform $\mathcal{D}(\bm{x})$ is defined as
\begin{equation*}
  \mathcal{D}(\bm{x})= \left\{
  \begin{array}{ll}
    0 & \quad \bm{x} \in \text{solid phase} \\
    \min\limits_{y \in \text{solid phase}} \rho(\bm{x},\bm{y}) & \quad \text{otherwise}
  \end{array}
\right.
\end{equation*}

Our algorithm for calculating $P(r)$ for a two-phase system $a$ is the
following:
\begin{algorithmic}[1]
  \Procedure{poresize}{$a, nbins$}
    \State $d \gets dist(a)$ \Comment{$dist(a)$ calculates
      $\mathcal{D}(\bm{x})$ for each element of $a$}
    \State $f(x) \gets x \ne 0$
    \State $d' \gets filter(f, d)$
    \State Bin values in $d'$ to $nbins$ bins to compute a histogram $H$
    \State \textbf{return} $H$
  \EndProcedure
\end{algorithmic}
Algorithmic efficiency is $O(n)$ where $n$ is the number of elements in the
input array.

@@

\subsection{Phase cross-corellation function}

Phase cross-correlation function $\rho_{ij}(\bm{r})$ can be defined as
\begin{equation*}
  \rho_{ij}(\bm{r}) = \langle I^{(i)}(\bm{x}) I^{(j)}(\bm{x} + \bm{r}) \rangle
\end{equation*}
which can be rewritten as \cref{eq:s2gen} with $\chi(\bm{x_1}, \bm{x_2})$
defined as follows:
\begin{equation*}
  \chi(\bm{x_1}, \bm{x_2}) = I^{(i)}(\bm{x}) I^{(j)}(\bm{x})
\end{equation*}
Now to calculate cross-correlation function we can use \code{s2sep} function
described in \cref{sec:s2}.

>>>>

The aim of this paper is to establish a general open-source solution allowing to
compute all classical correlation functions. This solution should be well
documented with all computational algorithms explained, it should also leverage
on recent advances in programming high-intensity computations and utilize both
CPU and GPU power. Our answer in the form of \code{CorrelationFunctions.jl}
package allows to compute numerous CFs easily on any operating system. The
manuscript describing the package is organized as follows: in \cref{sec:math} we
introduce a reader to the most common and useful correlation functions,
\cref{sec:map} presents all algorithmic details for computing full CFs maps (map
method, full 2-point statistics for a given image). In \cref{sec:scan} we
discuss the second method to calculate correlation functions which is called the
scan approach. In \cref{sec:verification} we verify the all computations based on
known analytical solutions. \cref{sec:efficiency} deals with computational
efficiency of our code, we compare computational times for 2D and 3D images of
different sizes for scanning and map methods as evaluated using either CPU or
GPU. In \cref{sec:examples} some application examples are provided, including
CFs evaluation for multi-phase materials. Finally, \cref{sec:summary} summarizes
all results and outlines possible future improvements.

@@

One of basic correlation functions is a lineal path function which is defined as
follows for homogenous media:
\begin{definition}
  Lineal path function $L_2^{(i)}(\bm{r})$ equals to probability that a
  vector $\bm{r}$ lies wholly in phase $i$ when randomly thrown into the
  sample.
\end{definition}
If a sample is isotropic a vector $\bm{r}$ can be replaced with its length
$r$ and the definition becomes:
\begin{definition}
  Lineal path function (for isotropic media) $L_2^{(i)}(r)$ equals to
  probability that a line segment of length $r$ lies wholly in phase $i$ when
  randomly thrown into the sample.
\end{definition}

Another function of big importance is a two-point correlation function $S_2$. A
definition for homogenous media is as follows:
\begin{definition}
  Two point function $S_2^{(i)}(\bm{r})$ equals to probability that both ends
  of a vector $\bm{r}$ lie in phase $i$ when the vector is randomly thrown
  into the sample.
\end{definition}
For isotopic media a vector $\bm{r}$ can be again reduced to a scalar value
$r$ (i.e. a correlation function depends only on length of a vector and not on
its direction.) Mathematically this definition can be expressed with the
following equation:
\begin{equation}
  S_2^{(i)}(\bm{r}) = \langle I^{(i)}(\bm{x}) I^{(i)}(\bm{x} +
  \bm{r}) \rangle
  \label{eq:s2-def}
\end{equation}
where $I^{(i)}$ is an indicator function for a set of all points belonging to
the phase $i$ and $\langle \cdots \rangle$ is an average over all xs.

A function closely related to $S_2$ is phase cross-correlation function
$\rho_{ij}$ which is defined with the following equation:
\begin{equation}
  \rho_{ij}(\bm{r}) = \langle I^{(i)}(\bm{x}) I^{(j)}(\bm{x} +
  \bm{r}) \rangle
  \label{eq:cross-def}
\end{equation}

The next function is a cluster function $C_2$.
\begin{definition}
  Cluster function $C_2^{(i)}(\bm{r})$ equals to probability that both
  ends of a vector $\bm{r}$ lie in the same cluster when the vector is
  randomly thrown into the sample. A cluster is a set of points of the same
  phase $i$ where any to points can be connected by a path lying entirely in
  that phase.
\end{definition}

The next two functions describe surface of a sample. These functions are
surface-surface and surface-void functions. For homogenous media their
definitions are:
\begin{definition}
  Surface-surface function $F_{ss}^{(i)}(\bm{r})$ equals to probability that
  both ends of a vector $\bm{r}$ touch the interface of phase $i$ when the
  vector is randomly thrown into the sample.
\end{definition}
and
\begin{definition}
  Surface-void function $F_{sv}^{(i)}(\bm{r})$ equals to probability that
  one end of a vector $\bm{r}$ touches the interface of phase $i$ and the
  other end touches the void phase when the vector is randomly thrown into the
  sample.
\end{definition}
If we introduce an interface indicator function
\begin{equation*}
  M^{(i)}(\bm{r}) = | \nabla I^{(i)}(\bm{r}) |
\end{equation*}
we can define these functions mathematically as
\begin{align}
  F_{ss}^{(i)}(\bm{r}) &= \langle M^{(i)}(\bm{x}) M^{(i)}(\bm{x} +
  \bm{r}) \rangle \label{eq:fss-def} \\
  F_{sv}^{(i)}(\bm{r}) &= \langle M^{(i)}(\bm{x}) I^{(void)}(\bm{x}
  + \bm{r}) \rangle \label{eq:fsv-def}
\end{align}
As always, there are versions of these functions applicable to scalar argument
for isotropic media.

@@

Cluster function can be calculated in the similar manner. Firstly, we apply an
indicator function to an input $A$: $A^{(i)} \xleftarrow{I^{(i)}} A$. When we
label connected components in $A^{(i)}$ using one of many popular algorithms for
this purpose \textcolor{red}{cite something}. This algorithm must take into
account if the function is calculated in periodic mode or not. Then $C_2$
function is a sum of $S_2$ functions calculated for all clusters
separately. Algorithmic complexity of this implementation is $O(MC \log M)$
where $M$ is the number of elements in the input and $C$ is the number of
clusters. The algorithm is as follows:
\begin{algorithmic}[1]
  \Procedure{$C_2$}{$A, phase, periodic$}
  \State $A^{(phase)} \gets I^{(phase)} (A)$
  \State $CC \gets label\_components(A^{(phase)}, periodic)$
  \State $N \gets maximum(CC)$ \Comment Number of connected components
  \State \textbf{return} $\sum\limits_{n=1}^N S_2(CC, n, periodic)$
  \EndProcedure
\end{algorithmic}

@@

All these algorithms contain parallelizable operations such as fast Fourier
transform, connected components labeling and image filtering, hence they are
perfectly suited for execution on GPUs.

\section{The scan approach.}
\label{sec:scan}
All algorithms described above are executed on the whole input and hence require
a lot of memory. The scan approach deals with memory limitations at the cost of
higher execution time. In this approach we select a list of predefined
directions and cut all one dimensional slices from the input along those
directions. Then for each direction we use an appropriate algorithm in
\cref{sec:map} to calculate a correlation function just for that slice. When all
slices are processed we calculate the result using the formula:
\begin{equation*}
  F^{\bm{d}}(r) = \frac{\sum\limits_i F^{\bm{d}}_i(r) N_i(r)}{\sum\limits_i N_i(r)}
\end{equation*}
Here $F^{\bm{d}}(r)$ is a correlation function computed with respect to a
direction $\bm{d}$, $F^{\bm{d}}_i(r)$ is a ``partial'' correlation
function calculated for a $i$-th slice and $N_i(r)$ is a total number of trials
for a correlation length $r$ and slice $i$.

\textcolor{red}{Lineal path function and chord length function}

Another function which does not use this approach but is in the
\code{Directional} module for historical reasons is the pore size function
$P(r)$. Remember that pore size function is a probability distribution
function. In practice it is better represented as a histogram of pore size
fractions, so all bars of the histogram sum to 1. In our approach we calculate
this histogram using distance transform. For each point $\bm{x}$ distance
transform $\mathcal{D}(\bm{x})$ is defined as
\begin{equation*}
  \mathcal{D}(\bm{x})= \left\{
  \begin{array}{ll}
    0 & \quad \bm{x} \in \text{solid phase} \\
    \min\limits_{y \in \text{solid phase}} \rho(\bm{x},\bm{y}) & \quad \text{otherwise}
  \end{array}
\right.
\end{equation*}
Distance transform has many implementations \textcolor{red}{cite something}. Our
algorithm for calculating $P(r)$ is then:
\begin{algorithmic}[1]
  \Procedure{$P$}{$A, nbins$}
    \State $D \gets \mathcal{D}(A)$ \Comment{Apply $\mathcal{D}$ to each element of $A$}
    \State $D' = \{\forall x \in D: x \ne 0\}$
    \State Bin values in $D'$ to $nbins$ bins to compute a histogram $H$
    \State \textbf{return} $H$
  \EndProcedure
\end{algorithmic}
Algorithmic efficiency is $O(n)$ where $n$ is the number of elements in the
input array.

@@
